### 2020-05-28 writeup/transfer.tex
sec 4

Based on Theorem \ref{thm_main_informal}, we provide precise interpretations of many empirical phenomena in multi-task learning.
We explain three contributing causes of negative transfer, including model dissimilarity, large source data size and covariate shift.
On the positive side, we describe two benefits of multi-task learning, including improving labeled data efficiency and de-noising via less noisy source labels.

@@

It is well-known since the seminal work of Caruana \cite{C97} that how well multi-task learning performs depends on task relatedness.
As a warmup, we first describe an example that rigorously formalizes the belief using our setup.

@@

\subsection{Implications on the Benefits of Multi-task Learning}\label{sec_benefit}

\textbf{De-noising by Transferring from a Source Task}
We further show that multi-task learning is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
We derive a sharp transition from positive to negative transfer as a parameter of $\sigma_1^2$.

\begin{proposition}\label{prop_var_transition}
	In the simplified model with $d$ being fixed but $\sigma_1$ varies, assume that $\rho_1 > 50$ is a fixed constant and $d^2 < \frac {\sigma_2^2} {2p} \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -\frac32 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
The details can be found in Appendix \ref{app_proof_data}.

\textbf{Improving Labeled Data Efficiency}
We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized by a metric called the data efficiency ratio as follows.
Given several tasks, let $\alpha^{\star}$ be the largest factors such that the total number of labeled datapoints needed for solving all the tasks can be reduced by an $\alpha^{\star}$ factor (compared to training independently) while keeping the performance nearly the same.
More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
If we only use $\alpha n_i$ datapoints from every task to train the multi-task learning estimator $\hat{\beta}(\alpha)$, then $\alpha \in (0, 1)$ will be the smallest number such that
\[ \alpha^{\star} \define\argmin_{\alpha\in(0, 1)} ~~ \te_1(\hat{\beta}(\alpha)) + \te_2(\hat{\beta}(\alpha))\le \te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL}). \]
We quantify the data efficiency ratio of $\hat{\beta}_t^{\MTL}$ for the simplified model as follows.

\begin{proposition}\label{prop_data_efficiency}
	In the simplified model, assume that $\rho_1 = \rho_2 = \rho \ge 200$ and $d^2 < {8\sigma^2} /{(3p \rho)}$.
	Then the data efficiency ratio is at most $\frac{1}{2\rho} + \frac{\sigma^2}{2\sigma^2 - 3p d^2 \rho / 4}$.
\end{proposition}
Note that we have stated the result assuming that $\rho_1 = \rho_2$.
Similar results can also be obtained when they are different.
We omit the details.
The proof of Proposition \ref{prop_data_efficiency} can be found in Appendix \ref{app_proof_data}.

>>>>

Based on Theorem \ref{thm_main_informal}, we provide precise interpretations of many empirical phenomena in multi-task learning.
In Section \ref{sec_similarity}, we explain three contributing causes of negative transfer, including model dissimilarity, large source data size and covariate shift.
In Section \ref{sec_benefit}, we describe two implications for positive transfer, including labeled data efficiency and de-noising.

@@

It is well-known since the seminal work of Caruana \cite{C97} that how well multi-task learning performs depends on task relatedness.
As a warmup, we first describe an example that rigorously formalizes the connection using our setup.

@@

\subsection{Implications for Positive Transfer}\label{sec_benefit}

On the positive side, we describe two implications for positive transfer, assuming that $\norm{\beta_1 - \beta_2}^2$ is small in the isotropic model.

\textbf{Labeled data efficiency.}
We use our tools to explain a key result of taskonomy \cite{ZSSGM18}, which shows that by learning from multiple related tasks, one can reduce the amount of labeled data from each task.
This is formalized by a metric called the data efficiency ratio as follows.

Given several tasks, let $\alpha^{\star}$ be the largest factors such that the total number of labeled datapoints needed for solving all tasks can be reduced by an $\alpha^{\star}$ factor (compared to training independently) while keeping the same performance.
%More precisely, suppose we have $n_i$ datapoints for each task, for $i= 1, 2$.
For $i = 1, 2$, let $\hat{\beta}_i^{\MTL}(\alpha)$ denote the estimator trained using only $\alpha n_i$ datapoints from every task.
The data efficiency ratio is defined as
\[ \alpha^{\star} \define\argmin_{\alpha\in(0, 1)} ~~
		\te_1(\hat{\beta}_1^{\MTL}(\alpha)) + \te_2(\hat{\beta}_2^{\MTL}(\alpha))
		\le \te_1(\hat{\beta}_1^{\STL}) + \te_2(\hat{\beta}_2^{\STL}). \]
We quantify the data efficiency ratio of the isotropic model as follows.

\begin{proposition}\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1$ and $\rho_2$, $pd^2 < {8\sigma^2} /{(3 \rho)}$.
	Then the data efficiency ratio is at most $\frac{1}{2\rho} + \frac{\sigma^2}{2\sigma^2 - 3p d^2 \rho / 4}$.
\end{proposition}
The proof of Proposition \ref{prop_data_efficiency} can be found in Appendix \ref{app_proof_data}.

\textbf{Labeled data de-noising.}
We further show that multi-task learning is particular powerful when the labeled data of the source task is less noisy compared to the target task.
Consider a more general setting of Proposition \ref{prop_dist_transition} where the noise level $\sigma_1$ of task $1$ differs from the noise level $\sigma_2$ of task $2$.
We extend Proposition \ref{prop_dist_transition} as follows.

\begin{proposition}\label{prop_var_transition}
	In the isotropic model, assume that $\rho_1 > 40$ and $pd^2 < \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -2 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
The details can be found in Appendix \ref{app_proof_data}.
